---
title: "straw1"
author: "ma415"
format: 
  html:
    embed-resources: true
editor: visual
---

#URL=https://quickstats.nass.usda.gov/results/806E4FDF-C7DF-3DCC-8381-B220597B08EC

# Preparing data for analysis

## Introduction: foundations

Before we begin to work with the strawberry data, let's talk about how we will approach the work.

### Data cleaning and organization

Cleaning and organizing data for analysis is an essential skill for data scientists. Serious data analyses must be presented with the data on which the results depend. The credibility of data analysis and modelling depends on the care taken in data preparation and organization.

#### References

In their handbook ["An introduction to data cleaning with R" by Edwin de Jonge and Mark van der Loo](https://cran.r-project.org/doc/contrib/de_Jonge+van_der_Loo-Introduction_to_data_cleaning_with_R.pdf), de Jonge and van der Loo go into detail about specific data cleaning isssues and how to handle them in R.

["Problems, Methods, and Challenges in Comprehensive Data Cleansing" by Heiko Müller and Johann-Christoph Freytag](https://www.researchgate.net/profile/Heiko-Mueller/publication/228929938_Problems_methods_and_challenges_in_comprehensive_data_cleansing/links/09e415101b58541e2c000000/Problems-methods-and-challenges-in-comprehensive-data-cleansing.pdf) is a good companion to the de Jonge and van der Loo handbook, offering additional issues in their discussion.

### Attitudes

Mechanistic descriptions of data cleaning methods are insufficient.

#### Data is the product (or by-product) of purposeful human activity

Much of the data used in analysis accessed on local databases or online which may create the impression that the data have been carefully curated. Beware. Data are produced by people for a purpose, with a point-of-view, and at a time and location that may affect the data. The provenance and lineage of the data are meta data you should include when reporting analysis. Data collection is purposeful human activity with all of the risks and weaknesses that are part of any purposeful human activity.

#### Data is language

Data has meaning. Data can be included in sentences related to the meaning of the data. Cleaning and organizing data should be informed by the meaning the data convey and how that meaning relates to the research you are doing do achieve this important result.

-   Immerse yourself in the data. Put data into context.

-   Visualize the data to find problems, confirm your understandings, and plan your data organization. People do a bad job of seeing meaningful patterns in data but a good job of seeing patterns of all kinds when data are rendered as plots. As you product and show visualizations, ask your self and those who view your presentations, "what do you see?" and "what do you wonder?"

## Example: Strawberries

### Public information

[WHO says strawberries may not be so safe for you--2017March16](https://med.news.am/eng/news/13621/who-says-strawberries-may-not-be-so-safe-for-you.html)

[Pesticides + poison gases = cheap, year-round strawberries 2019March20](https://www.ewg.org/foodnews/strawberries.php)

[Multistate Outbreak of Hepatitis A Virus Infections Linked to Fresh Organic Strawberries-2022March5](https://www.cdc.gov/hepatitis/outbreaks/fresh-strawberries-2022/?CDC_AAref_Val=https://www.cdc.gov/hepatitis/outbreaks/2022/hav-contaminated-food/index.htm)

[Strawberry makes list of cancer-fighting foods-2023May31](https://issuu.com/mechlocal/docs/053123_mech_asf/s/25386339)

## What is the question?

-   Where they are grown? By whom?

-   Are they really loaded with carcinogenic poisons?

-   Are they really good for your health? Bad for your health?

-   Are organic strawberries carriers of deadly diseases?

-   When I go to the market should I buy conventional or organic strawberries?

## The data

The data set for this assignment has been selected from:

[strawberries 2025feb24](https://quickstats.nass.usda.gov/results/806E4FDF-C7DF-3DCC-8381-B220597B08EC)

<!-- and has been stored on the blackboard as strawberries25_v3.csv. -->

## USDA NASS

```{r}
#| label: load libraries
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
```

## Read the file

```{r}
#| label: read data - glimpse 

data <- read_csv("strawb_mar6.csv", col_names = TRUE)

glimpse(data)
```

Some of original data is str, so we change it to num.

```{r}

##transfer to numerical

data <- data %>%
  mutate(
    Value = as.numeric(
      str_replace_all(
        str_replace_all(Value, ",", ""),
        "^(\\(D\\)|\\(NA\\)|\\(Z\\))$", NA_character_
      )
    ),
    `CV (%)` = as.numeric(
      str_replace_all(
        str_replace_all(`CV (%)`, ",", ""),
        "^(\\(D\\)|\\(NA\\)|\\(Z\\))$", NA_character_
      )
    )
  )

summary(data)


```

Function to eliminate the single value row.

```{r}
# eliminate single value column
drop_one_col <- function(data) {
  judge_c <- sapply(data, function(col) {
    if (is.list(col)) col <- unlist(col)
    length(unique(col)) == 1
  })
  return(data[, !judge_c, drop = FALSE])
}


data <- drop_one_col(data)

#delete empty col
data <- data[, !apply(is.na(data), 2, all)]


```

split data

```{r}

#seperate

data <- data |>
  separate_wider_delim(  cols = `Data Item`,
                         delim = " - ",
                         names = c("Fruit",
                                 "Category"),
                         too_many = "error",
                         too_few = "align_start"
                       )

data <- data |>
  separate_wider_delim(  cols = Fruit,
                         delim = ", ",
                         names = c("Fruit",
                                 "ORGANIC",
                                 "Organic detail"),
                         
                         too_many = "error",
                         too_few = "align_start"
                       )

data$Commodity <- NULL



data <- data |>
  separate_wider_delim(  cols = `Domain Category`,
                         delim = ": ",
                         names = c("Chemical",
                                 "Amount"),
                         
                         too_many = "error",
                         too_few = "align_start"
                       )

data$Domain <- NULL
```

We wanna to discuss about organic and conventional strawberry. So we filter the data with organic and chemical.

```{r}

# filter by chemical

chemical_data <- data |>
  filter(str_detect(Chemical, "^CHEMICAL"))

chemical_data <- chemical_data |>
  separate_wider_delim(  cols = Chemical,
                         delim = ", ",
                         names = c("null",
                                 "Chemical_Genre"),
                         
                         too_many = "error",
                         too_few = "align_start"
                       )

chemical_data <- chemical_data |> drop_one_col()

```

```{r}

#filter by organic

fruit_data <- data |>
  filter(Fruit == "STRAWBERRIES")

```

**Plot**

The red plot shows the total area of strawberry planting. We can find Cali and Florida plant most of the strawberry in area scale. Blue one is focusing on organic data.

```{r}
#fine the area data from fruit_data
library(ggplot2)
fruit_data$Category
acre <- fruit_data %>%
  filter(Category == "ACRES HARVESTED")

acre$Organic <- ifelse(is.na(acre$ORGANIC),"Conventional","ORGANIC")

#NA for converntional 
state_summary <- acre %>%
  group_by(State) %>%
  summarise(Total_Acres = sum(Value, na.rm = TRUE)) %>%
  arrange(desc(Total_Acres))

print(state_summary)
kable(state_summary, caption = "State-wise Total Acres Summary")%>%
  kable_styling(font_size = 14, bootstrap_options = c("striped", "hover", "condensed", "responsive"))
acre$Organic <- ifelse(acre$ORGANIC == "ORGANIC",1,0)

# visualize of total area of planting starberry
ggplot(state_summary, aes(x = reorder(State, Total_Acres), y = Total_Acres)) +
  geom_col(fill = "red") +
  coord_flip() +
  theme_minimal()

organic_summary <- acre %>%
  group_by(Organic) %>%
  summarise(Total_Acres = sum(Value, na.rm = TRUE)) %>%
  arrange(desc(Total_Acres))
print(organic_summary)

# visualize of organic labeled area
acre_organic <- acre |>
  filter(ORGANIC=="ORGANIC")

ggplot(acre_organic, aes(x = reorder(State, Value), y = Value)) +
    geom_col(fill = "blue") +
  coord_flip() +
  theme_minimal()
```

```{r}

fruit_data$ORGANIC
organic_data <- filter(fruit_data,ORGANIC=="ORGANIC")

```

We want to find if there are some harmful chemical used in planting. So we match toxic list with the chemical data. The result is that we find the MALATHION is widely used in California and Florida. Then we make a graph for 2021 and 2023 usage.

```{r}

#matching the toxic list
list <- "CAPTAFOL|DIAZINON|MALATHION|Dibromoethane|Glyphosate|DDT"

row_indices <- which(grepl(list, chemical_data$Amount, ignore.case = TRUE))

matches <- chemical_data[row_indices, ]

matches$OriginalRow <- row_indices


match_clean<-na.omit(matches)
match_clean<-match_clean %>%
  filter(Category=="APPLICATIONS, MEASURED IN LB / ACRE / YEAR, AVG")

head(match_clean)


library(ggplot2)

barplot(match_clean$Value)

yearly_summary <- match_clean %>%
  group_by(Year) %>%
  summarise(Total_Value = Value)


ggplot(yearly_summary, aes(x = factor(Year), y = Total_Value, fill = factor(Year))) +
  geom_bar(stat = "identity") +
  labs(title = "Yearly Value Distribution", x = "Year", y = "MALATHION USE") +
  theme_minimal() +
  scale_fill_manual(values = c("2021" = "steelblue", "2023" = "orange")) +
  theme(legend.position = "none")  

```

We want to find the chemicals that are used the most per acre per year on average.

Most of chemical in the graph is toxic. Especially for the Top 3, **Chloropicrin**, **Metam-potassium** and **Dichloropropen**e, they are highly toxic.

```{r}
#find the most amount of chemical 

chemical_data2 <- chemical_data |>
  filter(Category == "APPLICATIONS, MEASURED IN LB / ACRE / YEAR, AVG")

result <- chemical_data2 %>%
  group_by(Amount) %>%
  summarise(AvgValue = mean(Value, na.rm = TRUE))
print(result)

result_no_zero <- result %>%
  filter(AvgValue != 0)|>
  arrange(desc(AvgValue))
head(result_no_zero)
#top 10
result_clean<-head(result_no_zero,10)

kable(result_clean)%>%
  kable_styling(font_size = 14, bootstrap_options = c("striped", "hover", "condensed", "responsive"))

ggplot(result_clean, aes(x=reorder(Amount,AvgValue), y=AvgValue))+
  geom_bar(stat = "identity", fill = "green", width = 0.8)+
  theme_minimal()+
  labs(title="Chemical Amount",
       x="Chemical",
       y="Average Value")+
  theme(axis.text.y = element_text(size = 10))+
  geom_text(aes(label = round(AvgValue, 1)),
            angle = 0, hjust = 0, vjust = 0.4, size = 2)+
  coord_flip()


  
```

We also want to find the most widely used chemicals.

Notably, several widely used chemicals such as **Captan**, **Fludioxonil**, and **Chloropicrin** are associated with significant health risks. Captan, for example, is classified as a probable carcinogen under California’s Proposition 65, and Chloropicrin is a highly toxic soil fumigant known to cause respiratory irritation.

```{r}
chemical_data3 <- chemical_data |>
  filter(Category == "TREATED, MEASURED IN PCT OF AREA BEARING, AVG")

result2 <- chemical_data3 |>
  group_by(Amount) |>
  summarise(AvgValue = mean(Value, na.rm = TRUE))
print(result2)

result_no_zero2 <- result2 %>%
  filter(AvgValue != 0)|>
  arrange(desc(AvgValue))

result_clean2<-head(result_no_zero2,20)


ggplot(result_clean2, aes(x=reorder(Amount,AvgValue), y=AvgValue))+
  geom_bar(stat = "identity", fill = "purple", width = 0.8)+
  theme_minimal()+
  labs(title="Area percentage",
       x="Chemical",
       y="Area Percentage")+
  theme(axis.text.y = element_text(size = 10))+
  geom_text(aes(label = round(AvgValue, 1)),
            angle = 0, hjust = 0, vjust = 0.4, size = 2)+
  coord_flip()
```

Then, I try to find the organic sales in fresh market. In the graph, Florida provides most of organic strawberry sales' revenue. But, unfortunately, we are lack of data from California, the biggest planting area of strawberry.

```{r}
#organic sales in fresh market
organic_freshmarket <- organic_data |> 
  filter(`Organic detail`== "FRESH MARKET", Category == "SALES, MEASURED IN $") |>
  drop_one_col()
head(organic_freshmarket)

data2<- organic_freshmarket %>%
  filter(!is.na(Value)) %>%
  select(1:3)

#barplot 
ggplot(data2, aes(x = State, y = Value, fill = State)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  labs(title = "State-wise Value Distribution", x = "State", y = "Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

kable(data2, caption = "State-wise Value Table", align = "c") %>%
  kable_styling(
    font_size = 14, 
    bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
    full_width = FALSE
  ) %>%
  column_spec(1, bold = TRUE) 

```
